---
title: "CSA Comparison"
author: "Sam Neylon"
date: '2022-07-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(fs)
library(stringr)

Sys.sleep(2)

```

# Notes

Please create a new .Rmd file for each data pull, by saving this template with a name similar to the excel files which are exported. This will allow us to keep track of how each file was created, even though it will create a lot of markdown files.

## Changelog

### (08-02-2022)

Modifed cleaning function:

1. Added NAICS_4 and NAICS_5 variables
2. Added MSA_ag which flags MSA aggregates
3. Added NYC_Boro which flags NYC boroughs using FIPS codes

*CHECK* Everything seems to be working - the proper NAICS levels are there, the MSA codes are flagged, and the boroughs are flagged

#### EXPORT

08-02-22 Exports:

output/CSA_comparison_LifeSci.xlsx
output/CSA_comparison.RDS

# Functions

## get_dir

```{r eval=T}

# Assemble File Names
  # I have commented out a thing which saved each filepath file outside the function, but it may be useful for future projects.

get_dir <- function(fips){
  df <- dir_ls(path = "D:/QCEW/by_area/2007_2021", regexp = fips, recurse = TRUE)
  #assign(paste("c",fips, sep="_"), df, envir = .GlobalEnv)
}

# read_csv using file names

county_map <- function(fs_list){
  map_df(fs_list, read_csv, col_types = cols(
  area_fips = col_character(),
  own_code = col_double(),
  industry_code = col_character(),
  agglvl_code = col_double(),
  size_code = col_double(),
  year = col_double(),
  qtr = col_character(),
  disclosure_code = col_character(),
  area_title = col_character(),
  own_title = col_character(),
  industry_title = col_character(),
  agglvl_title = col_character(),
  size_title = col_character(),
  annual_avg_estabs_count = col_double(),
  annual_avg_emplvl = col_double(),
  total_annual_wages = col_double(),
  taxable_annual_wages = col_double(),
  annual_contributions = col_double(),
  annual_avg_wkly_wage = col_double(),
  avg_annual_pay = col_double(),
  lq_disclosure_code = col_character(),
  lq_annual_avg_estabs_count = col_double(),
  lq_annual_avg_emplvl = col_double(),
  lq_total_annual_wages = col_double(),
  lq_taxable_annual_wages = col_double(),
  lq_annual_contributions = col_double(),
  lq_annual_avg_wkly_wage = col_double(),
  lq_avg_annual_pay = col_double(),
  oty_disclosure_code = col_character(),
  oty_annual_avg_estabs_count_chg = col_double(),
  oty_annual_avg_estabs_count_pct_chg = col_double(),
  oty_annual_avg_emplvl_chg = col_double(),
  oty_annual_avg_emplvl_pct_chg = col_double(),
  oty_total_annual_wages_chg = col_double(),
  oty_total_annual_wages_pct_chg = col_double(),
  oty_taxable_annual_wages_chg = col_double(),
  oty_taxable_annual_wages_pct_chg = col_double(),
  oty_annual_contributions_chg = col_double(),
  oty_annual_contributions_pct_chg = col_double(),
  oty_annual_avg_wkly_wage_chg = col_double(),
  oty_annual_avg_wkly_wage_pct_chg = col_double(),
  oty_avg_annual_pay_chg = col_double(),
  oty_avg_annual_pay_pct_chg = col_double())
)
}

```

## get_dir_year

```{r eval=T}

get_dir_year <- function(fips, year_string){
  df <- dir_ls(path = year_string, regexp = fips, recurse = TRUE)
  #assign(paste("c",fips, sep="_"), df, envir = .GlobalEnv)
}

```

## All years Function

This function will get data for all the years in the directory "D:/QCEW/by_area/2007_2021" (back to 2007)

```{r eval=T}

# By FIPS

qcew_FIPS <- function(fips_list){
  # Get file paths using get_dir(), mapped over list of FIPS
  fips_dirs <- map(fips_list, get_dir)
  # Turn filepaths into dataframes, mapping county_map over list of file paths, which takes file paths and feeds them into read_csv
  county_df <- map(fips_dirs, county_map)
  # Do a simple bind_rows to turn your list (a list of dataframes) into one large data frame
  df <- map_dfr(county_df, bind_rows)
}


```

## By Year Function

This is the one we will use the most. It takes a list of years and list of FIPS and turns them into a dataframe.

```{r eval=T}

qcew_FIPS_year <- function(fips_list, qcew_year){
  k <- str_c("D:/QCEW/by_area/2007_2021/", qcew_year, "_annual_by_area/")
  # Get file paths using get_dir(), mapped over list of FIPS
  fips_dirs <- map(fips_list, ~ get_dir_year(.x, year_string = k))
  # Turn filepaths into dataframes, mapping county_map over list of file paths, which takes file paths and feeds them into read_csv
  county_df <- map(fips_dirs, county_map)
  # Do a simple bind_rows to turn your list (a list of dataframes) into one large data frame
  df <- map_dfr(county_df, bind_rows)
}

```

## Into List Function

This function turns a column of FIPS into a list.

```{r eval=T}

into_fipsList <- function(code_list){
  # Select FIPS Column
  df <- code_list %>% select(fips)
  # Turn this vector into a list with a separate item for each row.
  list_data <- split(df,seq(nrow(df)))  
}


```

## Cleaning Function

```{r eval=T}

NAICS_clean <- function(df){
df <- df %>% 
  mutate(NAICS_lvl = if_else((str_length(industry_code)==2), 2, 
                                     if_else((str_length(industry_code)==3), 3, 
                                             if_else((str_length(industry_code)==4), 4, 
                                                     if_else((str_length(industry_code)==5), 5, 6))))) %>% 
  mutate(NAICS_lvl = ifelse((str_detect(industry_code, "-")), 2, NAICS_lvl),
         dash_code = ifelse((str_detect(industry_code, "-")), 1, 0))

# NAICS substring

df <- df %>% mutate(NAICS_1 = str_sub(industry_code, 1, 1),
                              NAICS_2 = str_sub(industry_code, 1, 2),
                              NAICS_3 = ifelse(NAICS_lvl>=3, str_sub(industry_code, 1, 3), NA_character_),
                    NAICS_4 = ifelse(NAICS_lvl>=4, str_sub(industry_code, 1, 4), NA_character_),
                    NAICS_5 = ifelse(NAICS_lvl>=5, str_sub(industry_code, 1, 5), NA_character_)) %>% 
  mutate(NAICS_2 = ifelse(dash_code==1, industry_code, NAICS_2),
         NAICS_3 = ifelse(dash_code==1, NA_character_, NAICS_3))

df <- df %>% relocate(c("NAICS_lvl","NAICS_1","NAICS_2","NAICS_3","NAICS_4","NAICS_5"), .after = industry_code)
df <- df %>% 
  mutate(NAICS_17 = ifelse(year >= 2017, industry_code, NA_character_),
         NAICS_12 = ifelse(between(year, 2012, 2016), industry_code, NA_character_),
         NAICS_07 = ifelse(between(year, 2007, 2011), industry_code, NA_character_),
         MSA_ag = ifelse(str_detect(area_fips, "C"), 1, 0),
         NYC_Boro = ifelse(area_fips %in% c("36085","36081","36061","36047","36005"), 1, 0)) 
}

```

# Data Setup

## FIPS List

```{r eval=T}

all_fips <- read_csv(here("data/FIPS_all_counties.csv"))

all_fips <- all_fips %>% mutate(fips = ifelse(str_length(FIPS)==4, str_c("0",FIPS), FIPS))

```

## 2017 NAICS

```{r eval=T}

NAICS_titles <- read_csv(here("data/naics_2017.csv"))

# NAICS_titles <- NAICS_titles %>% select(industry_code, industry_title)

```


# Counties Template

Filter a list of counties by a larger geographic unit (i.e. MSA), and pull data for that list.

## Year/Geo Setup Block

```{r eval = F}

# Geographic Unit

geo_code <- "MSA or other Geographic CODE HERE"

# 1 year

years_list <- list("2020")

# Multiple Years

# years_list <- as.list(seq(2007, 2021, by = 1))

```

## Filter Counties

### Using geo_code

IMPORTANT NOTE: MUST CHANGE THE NAME OF FILTER VARIABLE IF IT ISN'T MSA

```{r eval=F}

# QCEW MSA Code

df_fips <- all_fips %>% filter(`QCEW MSA Code` == geo_code)

df_fips <- into_fipsList(df_fips)

```

### Using list of FIPS

This is if you want to put a custom list of FIPS into a csv, import it, and use that.

```{r eval=F}

df_fips <- read_csv(here("data/NAME OF CUSTOM FIPS CSV"))

df_fips <- into_fipsList(df_fips)

```

## Import and Clean

```{r eval=F}

# Import

df_1 <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = df_fips))

df_1 <- map_dfr(df_1, bind_rows)

rm(df_fips)

# Clean

df_1 <- NAICS_clean(df_1)

```

## Sum by Geography

The code above gives you data by county. This code summarizes all the counties, giving you a total employment for the whole geographic area (or list of FIPS). This is equivalent to the numbers you get by filtering for a QCEW MSA code (e.g. C4186 San Francisco-Oakland-Hayward).

This code summarizes by own_code==5 (private industry), otherwise all employment and establishments in government will be added to the private industry counts. You may be able to group by "own_code" to get a list of industry by own code sums.

```{r eval=F}

df_1_sum <- df_1 %>% 
  filter(own_code == 5) %>% 
  group_by(industry_code) %>% 
  summarise(ind_ests = sum(annual_avg_estabs_count),
            ind_emp = sum(annual_avg_emplvl))
df_1_sum <- df_1_sum %>% left_join(NAICS_titles, by = "industry_code") %>% 
  relocate(industry_title)

```

## Export

```{r eval=F}

openxlsx::write.xlsx(df_1_sum, here("output/FILE NAME HERE.xlsx"))

```


# Single Geography Template

This template allows you to get data for just one QCEW geographic unit, such as MSA (e.g. C4186 San Francisco-Oakland-Hayward) or State.

## Year/Geo Setup Block

```{r eval = F}

# Geographic Unit

df_fips <- list("MSA or other Geographic CODE HERE")

# 1 year

years_list <- list("2020")

# Multiple Years

# years_list <- as.list(seq(2007, 2021, by = 1))

```

## Import and Clean

```{r eval=F}

# Import

df_1 <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = df_fips))

df_1 <- map_dfr(df_1, bind_rows)

rm(df_fips)

# Clean

df_1 <- NAICS_clean(df_1)

```

## "Sum"

Because this is one geography, you don't really need to sum the industries. However, this creates a dataframe which can easily be joined to the County sum data and compared (see "Merge" section below).

```{r eval=F}

df_1_sum <- df_1 %>% 
  filter(own_code == 5,
         NAICS_lvl == 6) %>% 
  group_by(industry_code) %>% 
  summarise(ind_ests = sum(annual_avg_estabs_count),
            ind_emp = sum(annual_avg_emplvl))
df_1_sum <- df_1_sum %>% left_join(NAICS_titles, by = "industry_code") %>% 
  relocate(industry_title)

```

## Export

```{r eval=F}

openxlsx::write.xlsx(df_1, here("output/FILE NAME HERE.xlsx"))

```

# Merge Example

This example takes data based on a QCEW MSA code (SF_MSA_sum) and joins it with one based on a list of county FIPS within that MSA (SF_sum). This creates two ccolumns for industry establishments and employment for the first data frame (example below has MSA first), and two for the second (list of counties, referred to with "c_").

Since this merge loses the NAICS level codes I added in the cleaning stage, the code below adds in 1-5 digit NAICS codes.

The code below is designed for one year of data (or a single time frame).

```{r eval=F}

SF_comp <- full_join(SF_MSA_sum, SF_sum, by = "industry_code") %>% 
  mutate(MSA_ests = ind_ests.x,
         MSA_emp = ind_emp.x,
         C_ests = ind_ests.y,
         C_emp = ind_emp.y) %>% 
  select(industry_title.x, industry_code, MSA_ests, MSA_emp, C_ests, C_emp) %>% 
  mutate(MSA_C_ests = MSA_ests - C_ests,
         MSA_C_emp = MSA_emp - C_emp,
         NAICS_1 = str_sub(industry_code, 1, 1),
         NAICS_2 = str_sub(industry_code, 1, 2),
         NAICS_3 = str_sub(industry_code, 1, 3),
         NAICS_4 = str_sub(industry_code, 1, 4),
         NAICS_5 = str_sub(industry_code, 1, 5))

```

# CSA Comparison Section

Steps:

1. Get all CSA counties for three geographies...
  For all years? How about back to 2017, since I don't need a crosswalk.
2. Get one-code CSA's and MSA's for the geographies.
3. Create columns indicating:
  a. CSA column: CSA's for counties, or one-code for QCEW CSA
  b. MSA column: Same
4. Append all datasets
5. Filter industries?
5. Do analysis in Excel

## LifeSci NAICS

```{r eval=T}

LifeSci_NAICS <- read_csv(here("data/07-29_LifeSci_NAICS.csv"), col_types = cols(
  LifeSci_NAICS17 = col_character(),
  LifeSci_NAICS17_Title = col_character()))

```

## MSA CSA Columns

```{r eval=T}

csa_columns <- all_fips %>% select(fips, QCEW_MSA, QCEW_MSA_Code, QCEW_CSA, QCEW_CSA_TITLE )

```


## Boston CSA

*CS148*:	Boston-Worcester-Providence, MA-RI-NH-CT

### Year/Geo Setup Block

```{r eval=T}

# Geographic Unit

geo_code <- c("CS148")

# 1 year

# years_list <- list("2020")

# Multiple Years

years_list <- as.list(seq(2017, 2021, by = 1))

```

### Filter Counties

#### Using geo_code

```{r eval=T}

gc()

# QCEW MSA Code

BOS_fips <- all_fips %>% filter(QCEW_CSA == paste(geo_code))

BOS_fips <- into_fipsList(BOS_fips)

```

#### Import and Clean

```{r eval=T}

Sys.sleep(2)

# Import

BOS_counties <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = BOS_fips))

BOS_counties <- map_dfr(BOS_counties, bind_rows)

rm(BOS_fips)

# Clean

BOS_counties <- NAICS_clean(BOS_counties)

```

### BOS MSA

*C1446* Boston-Cambridge-Newton, MA-NH

#### Year/Geo Setup Block

```{r eval=T}

# Geographic Unit

df_fips <- list("C1446")

# 1 year

# years_list <- list("2020")

# Multiple Years

# years_list <- as.list(seq(2007, 2021, by = 1))

```

#### Import and Clean

```{r eval=T}

# Import

BOS_MSA <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = df_fips))

BOS_MSA <- map_dfr(BOS_MSA, bind_rows)

# Clean

BOS_MSA <- NAICS_clean(BOS_MSA)

```

## San Francisco CSA

*CS488*	San Jose-San Francisco-Oakland, CA

### Year/Geo Setup Block

```{r eval=T}

# Geographic Unit

geo_code <- c("CS488")

# 1 year

# years_list <- list("2020")

# Multiple Years

years_list <- as.list(seq(2017, 2021, by = 1))

```

### Filter Counties

#### Using geo_code

```{r eval=T}

# QCEW MSA Code

SF_fips <- all_fips %>% filter(QCEW_CSA == paste(geo_code))

SF_fips <- into_fipsList(SF_fips)

```

### Import and Clean

```{r eval=T}

Sys.sleep(2)

# Import

SF_counties <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = SF_fips))

SF_counties <- map_dfr(SF_counties, bind_rows)

rm(SF_fips)

# Clean

SF_counties <- NAICS_clean(SF_counties)

```

### SF MSA

*C4186* San Francisco-Oakland-Hayward, CA	

#### Year/Geo Setup Block

```{r eval=T}

# Geographic Unit

df_fips <- list("C4186")

# 1 year

# years_list <- list("2020")

# Multiple Years

# years_list <- as.list(seq(2007, 2021, by = 1))

```

#### Import and Clean

```{r eval=T}

# Import

SF_MSA <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = df_fips))

SF_MSA <- map_dfr(SF_MSA, bind_rows)

# Clean

SF_MSA <- NAICS_clean(SF_MSA)

```

## NYC CSA

*CS408*	New York-Newark, NY-NJ-CT-PA

### Year/Geo Setup Block

```{r eval=T}

# Geographic Unit

geo_code <- c("CS408")

# 1 year

# years_list <- list("2020")

# Multiple Years

years_list <- as.list(seq(2017, 2021, by = 1))

```

### Filter Counties

#### Using geo_code

```{r eval=T}

# QCEW MSA Code

NYC_fips <- all_fips %>% filter(QCEW_CSA == paste(geo_code))

NYC_fips <- into_fipsList(NYC_fips)

```

### Import and Clean

```{r eval=T}

Sys.sleep(2)

# Import

NYC_counties <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = NYC_fips))

NYC_counties <- map_dfr(NYC_counties, bind_rows)

rm(NYC_fips)

# Clean

NYC_counties <- NAICS_clean(NYC_counties)

```

### NYC MSA

*C3562* New York-Newark-Jersey City, NY-NJ-PA

#### Year/Geo Setup Block

```{r eval=T}

# Geographic Unit

df_fips <- list("C3562")

# 1 year

# years_list <- list("2020")

# Multiple Years

# years_list <- as.list(seq(2007, 2021, by = 1))

```

#### Import and Clean

```{r eval=T}

# Import

NYC_MSA <- map(years_list, ~ qcew_FIPS_year(qcew_year = .x, fips_list = df_fips))

NYC_MSA <- map_dfr(NYC_MSA, bind_rows)

# Clean

NYC_MSA <- NAICS_clean(NYC_MSA)

```

# Append

```{r eval=T}

CSA_comparison <- list(BOS_counties,
               BOS_MSA,
               SF_counties,
               SF_MSA,
               NYC_counties,
               NYC_MSA)
 
CSA_comparison <- bind_rows(CSA_comparison)

rm(BOS_counties,
               BOS_MSA,
               SF_counties,
               SF_MSA,
               NYC_counties,
               NYC_MSA)

```


## CSA MSA Columns

```{r eval=T}

CSA_comparison  <- CSA_comparison %>% left_join(csa_columns, by = c("area_fips" = "fips"))

```

# Export CSA_comparison

```{r eval=F}

openxlsx::write.xlsx(CSA_comparison, here("output/CSA_comparison.xlsx"))

```

# LifeSci Filter

```{r eval=T}

CSA_comparison_LifeSci <- CSA_comparison %>% semi_join(LifeSci_NAICS, by = c("industry_code" = "LifeSci_NAICS17"))

```

## Export LifeSci

```{r eval=F}

openxlsx::write.xlsx(CSA_comparison_LifeSci, here("output/CSA_comparison_LifeSci.xlsx"))

```

# Clear Sub-Tables

```{r eval=F}

rm(BOS_counties,
               BOS_MSA,
               SF_counties,
               SF_MSA,
               NYC_counties,
               NYC_MSA)

gc()

```


# CSA Summaries

NOTE: The first time I did this, I messed up. I have included multiple NAICS levels, so I was summing things more than once!

```{r eval=F}

# Summarize All LifeSci Industries by Area

LifeSci_Summary <- CSA_comparison_LifeSci %>% 
  filter(own_code == 5,
         NAICS_lvl == 6,
         year == 2021) %>% 
  group_by(area_fips) %>% 
  summarise(lifeSci_ests = sum(annual_avg_estabs_count),
            lifeSci_emp = sum(annual_avg_emplvl))

# Re-attach needed information

LifeSci_Summary <- CSA_comparison_LifeSci %>%
  select(area_fips, area_title, QCEW_MSA, QCEW_MSA_Code, QCEW_CSA, QCEW_CSA_TITLE) %>% 
  right_join(LifeSci_Summary, by = "area_fips") %>% 
  relocate(c("lifeSci_ests", "lifeSci_emp"), .after = area_title) %>% 
  distinct(.keep_all = TRUE)

```

## Export Summary

```{r eval=F}

openxlsx::write.xlsx(LifeSci_Summary, here("output/LifeSci_Summary.xlsx"))

```

## Just Counties

```{r eval=F}

LifeSci_Sum_Counties <- LifeSci_Summary %>% filter(!is.na(QCEW_CSA)) %>% 
  group_by(QCEW_CSA) %>% 
  mutate(ests_CSA = sum(lifeSci_ests), 
         est_CSApct = round((prop.table(lifeSci_ests) * 100), digits = 3),
         emp_CSA = sum(lifeSci_emp), 
         emp_CSApct = round((prop.table(lifeSci_emp) * 100), digits = 3)) %>% 
  ungroup() %>% 
  group_by(QCEW_MSA_Code) %>% 
  mutate(ests_MSA = sum(lifeSci_ests), 
         est_MSApct = round((prop.table(lifeSci_ests) * 100), digits = 3),
         emp_MSA = sum(lifeSci_emp), 
         emp_MSApct = round((prop.table(lifeSci_emp) * 100), digits = 3)) %>% 
  ungroup()

```

## Export Summary

```{r eval=F}

openxlsx::write.xlsx(LifeSci_Sum_Counties, here("output/LifeSci_Sum_Counties.xlsx"))

```

# All LifeSci Ind

Keep the smaller industries within the areas.

```{r eval=F}

# Summarize All LifeSci Industries by Area

LifeSci_Summary_byInd <- CSA_comparison_LifeSci %>% 
  filter(own_code == 5,
         NAICS_lvl == 6,
         year == 2021) %>% 
  group_by(area_fips, industry_code) %>% 
  summarise(ind_ests = sum(annual_avg_estabs_count),
            ind_emp = sum(annual_avg_emplvl))

# Re-attach needed information

LifeSci_Summary_byInd <- CSA_comparison_LifeSci %>%
  select(area_fips, area_title, QCEW_MSA, QCEW_MSA_Code, QCEW_CSA, QCEW_CSA_TITLE) %>% 
  right_join(LifeSci_Summary_byInd, by = "area_fips") %>% 
  relocate(c("ind_ests", "ind_emp"), .after = area_title) %>% 
  left_join(NAICS_titles, by = "industry_code") %>% 
  relocate(industry_code, industry_title, .after = area_title) %>% 
  distinct(.keep_all = TRUE)

```

## Export Summary

```{r eval=F}

openxlsx::write.xlsx(LifeSci_Summary_byInd, here("output/LifeSci_Summary_byInd.xlsx"))

```

## Calculated Columns

```{r eval=F}

LifeSci_Sum_byIND <- LifeSci_Summary_byInd %>% filter(!is.na(QCEW_CSA)) %>% 
  group_by(area_fips) %>% 
  mutate(ests_county = sum(ind_ests), 
         est_countypct = round((prop.table(ind_ests) * 100), digits = 3),
         emp_county = sum(ind_emp), 
         emp_countypct = round((prop.table(ind_emp) * 100), digits = 3)) %>% 
  ungroup() %>% 
  group_by(QCEW_CSA) %>% 
  mutate(ests_CSA = sum(ind_ests), 
         est_CSApct = round((prop.table(ind_ests) * 100), digits = 3),
         emp_CSA = sum(ind_emp), 
         emp_CSApct = round((prop.table(ind_emp) * 100), digits = 3)) %>% 
  ungroup() %>% 
  group_by(QCEW_MSA_Code) %>% 
  mutate(ests_MSA = sum(ind_ests), 
         est_MSApct = round((prop.table(ind_ests) * 100), digits = 3),
         emp_MSA = sum(ind_emp), 
         emp_MSApct = round((prop.table(ind_emp) * 100), digits = 3)) %>% 
  ungroup()

```

## Export Calculated Columns

```{r eval=F}

openxlsx::write.xlsx(LifeSci_Sum_byIND, here("output/LifeSci_Sum_byIND.xlsx"))

```

# RDS

I want to save the master CSA_comparison file as an .rds, so I can use it in other code more easily.

```{r eval=F}

saveRDS(CSA_comparison, here("output/CSA_comparison.RDS"))

```

