---
title: "County Mapping"
author: "Sam Neylon"
date: '2022-07-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(data.table)
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(sf)

```

# Census API Key

```{r eval=F}

census_api_key("aa512886c5449a582d837da8d3a07af66a043fe5", install = TRUE)

```


# Getting County GIS Data

```{r eval=T}

counties_df <- get_decennial(geography = "county",
                       variables = "P001001",
                       geometry = T,
                       output = "wide")

```

# Merge with CSA LifeSci

## NOTE

I have loaded the data I need by running the CSA_comparison.Rmd file in the same R session. I am using a file which is created at the bottom of that file - summaries by county.

## Merge

```{r eval=TRUE}

map_LifeSci_Sum_Counties <- left_join(LifeSci_Sum_Counties, counties_df, by = c("area_fips" = "GEOID"))

```

## Write to Shapefile

```{r eval=F}

st_write(map_LifeSci_Sum_Counties, here("maps/map_LifeSci_Sum_Counties.shp"))

```


# OLD CODE

## Export County Geography

Wait, this won't work, since it is an "sf" object...

```{r eval=F}

write_csv(counties_df, here("output/county_shapefiles.csv"))

```

# Import County List

```{r eval=F}

FIPS_list <- c(42103,36119,36103,36087,36085,36081,36079,36071,36061,36059,36047,36027,36005,34039,34037,34035,34031,34029,34027,34025,34023,34019,34017,34013,34003,06081,06075,06041,06013,06001,33017,33015,25025,25023,25021,25017,25009)

```

# Import and Merge

```{r eval=F}

UWS_fips <- fread(here("data/FIPS_filter.csv"), 
             stringsAsFactors = F, 
             data.table = F, 
             colClasses=list(character=c("FIPS")))

```

## Merge

```{r eval=F}

UWS_df <- left_join(UWS_fips, counties_df, by = c("FIPS" = "GEOID"))

```

# Write to Shapefile

```{r eval=F}

st_write(UWS_df, "uws_fips.shp")

```


# Porter Code

```{r eval=F}

setwd("C:/Users/Sam/OneDrive - Cuny GradCenter/CUNY! (Cloud)/2020.2 Fall/Quant Methods/Lab 7")
require(data.table)
require(tidycensus)
require(ggplot2)
#census_api_key("aa512886c5449a582d837da8d3a07af66a043fe5")

census_data <- load_variables(2018, "acs5", cache=T)
View(census_data)
fwrite(census_data, "census_variables.csv")

vars <- c(tpop = 'P001001',
          medage = 'P013001',
          wpop = 'P003002',
          bpop = 'P003003',
          apop = 'P003005',
          hpop = 'P004003')
View(vars)

NY_df <- get_decennial(state = "ny", 
                       geography = "county",
                       variables = vars,
                       geometry = T,
                       output = "wide")
View(NY_df)

plot(NY_df)
plot(NY_df["tpop"])
plot(NY_df["bpop"])
plot(NY_df["medage"])

NY_df$wpct <- (NY_df$wpop / NY_df$tpop) *100
NY_df$bpct <- (NY_df$bpop / NY_df$tpop) *100
NY_df$apct <- (NY_df$apop / NY_df$tpop) *100
NY_df$hpct <- (NY_df$hpop / NY_df$tpop) *100

###Compute Simpson's Diversity Index **D = ((SUM n(n-1))/N(N-1))**
###Higher value indicates higher diversity

NY_df$race_div <- 1 - (((NY_df$wpop*(NY_df$wpop-1))+
                          (NY_df$bpop*(NY_df$bpop-1))+
                          (NY_df$hpop*(NY_df$hpop-1))+
                          (NY_df$apop*(NY_df$apop-1)))/
                         (NY_df$tpop*(NY_df$tpop-1)))

head(NY_df)
summary(NY_df$race_div)
plot(NY_df["race_div"])

rwj <- fread("rwj_rank.csv", 
             stringsAsFactors = F, 
             data.table = F, 
             colClasses=list(character=c("FIPS")))
head(rwj)
ny_rwj <- subset(rwj, State == "New York")
head(ny_rwj, n=62)

ny_rwj_df <- merge(NY_df, 
                   ny_rwj,
                   by.x = "GEOID",
                   by.y = "FIPS")

summary(ny_rwj_df)
summary(ny_rwj_df$QL.Rank)
ny_rwj_df$QL.Rank <- as.numeric(ny_rwj_df$QL.Rank)
summary(ny_rwj_df$QL.Rank)

### use ggplot to visualize data in map
map1 <- ggplot(ny_rwj_df) +
  geom_sf(aes(fill = cut_number(QL.Rank, 5)))
map1

map2 <- ggplot(ny_rwj_df, aes(fill = QL.Rank)) +
  geom_sf() +
  scale_fill_continuous(low = "#34E8EB", high = "#3D34EB") +
  #scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(line = element_blank(),                          
        axis.text=element_blank(),                       
        axis.title=element_blank(),                      
        panel.background = element_blank()) 
map2

summary(ny_rwj_df$QL.Rank)
ny_rwj_df$QL.Rank <- (ny_rwj_df$QL.Rank - 63) * -1

map3 <- ggplot(ny_rwj_df) +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(hjust = 0.5))  
map3

map4 <- ggplot(ny_rwj_df) +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5)) 
map4

map5 <- ggplot(ny_rwj_df) +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5)) 
map5

map6 <- ggplot(ny_rwj_df) +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen",
                        breaks = c(12, 24, 36, 48, 60),
                        name = "Quality of Life",
                        labels = c("low",
                                   "",
                                   "",
                                   "",
                                   "high")) +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5)) 
map6

map7 <- ggplot(ny_rwj_df) +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen",
                        breaks = c(3, 12, 24, 36, 48, 60),
                        name = "Quality\nof Life",
                        labels = c("low",
                                   "",
                                   "",
                                   "",
                                   "",
                                   "high")) +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) 
map7

names(ny_rwj_df)
summary(ny_rwj_df$race_div)

map8 <- ggplot(ny_rwj_df) +
  geom_sf(aes(fill = race_div)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen",
                        breaks = c(0.05, 0.1, 0.16, 0.31, 0.65),
                        name = "Racial\nDiversity",
                        labels = c("low",
                                   "",
                                   "",
                                   "",
                                   "high")) +
  ggtitle("County Level Racial Diversity") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) 
map8


ny_rwj_df$race_div_bin <- ifelse(ny_rwj_df$race_div > mean(ny_rwj_df$race_div),
                                 "High Diversity",
                                 "Low Diversity")
ny_rwj_df$Qual_Life_bin <- ifelse(ny_rwj_df$QL.Rank > mean(ny_rwj_df$QL.Rank),
                                  "High Quality of Life",
                                  "Low Quality of Life")

plot1 <- ggplot(ny_rwj_df, 
                     aes(x = race_div_bin, 
                         y = QL.Rank, 
                         group = race_div_bin, 
                         color = race_div_bin, 
                         fill = race_div_bin)) + 
  geom_boxplot() + 
  labs(title="Boxplots", 
       subtitle= "Quality of Life by Racial Diversity",
       y="Quality of Life", 
       x="Racial Diversity") +
  scale_colour_brewer(palette = "Set3") +
  theme(legend.position = "none")
plot1

require(RColorBrewer)
display.brewer.all()

plot2 <- ggplot(ny_rwj_df, 
                aes(x = Qual_Life_bin, 
                    y = race_div, 
                    group = Qual_Life_bin, 
                    color = Qual_Life_bin, 
                    fill = Qual_Life_bin)) + 
  geom_boxplot() + 
  labs(title="Boxplots", 
       subtitle= "Racial Diversity by Quality of Life",
       y="Racial Diversity", 
       x="Quality of Life") +
  scale_colour_brewer(palette = "Set3") +
  theme(legend.position = "none")
plot2

install.packages("ggpubr")
require(ggpubr)

full_plot <- ggarrange(map7,
                       map8,
                       plot1,
                       plot2,
                       ncol = 2,
                       nrow = 2)
full_plot


### Please submit the result of the lab script in 
### fulfillment of your assignment 7 requirement


loc <-  getwd()
require(sf)

```

