---
title: "UWS Mapping 1"
author: "Sam Neylon"
date: '2022-07-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(ggplot2)
library(sf)
library(ggmap)

# Google Map API

  # register_google(key = DELETED)

```


# Read Map Data

From CSA Comparisons

```{r eval=T}

CSA_map <- st_read(here("maps/CSA_comparison"))

```
# Boston

## Basemap

### Toner Basemap

See: https://cengel.github.io/R-spatial/mapping.html#adding-basemaps-with-ggmap

Example code:

* Philadelphia Lat 39.95258 and Lon is -75.16522

ph_basemap <- get_map(location=c(lon = -75.16522, lat = 39.95258), zoom=11, maptype = 'terrain-background', source = 'stamen')

CartoDB.Positron

*Boston*

42.348527, -71.110049

```{r eval=F}

BOS_basemap <- get_stamenmap(location=c(lon = -71.110049, lat = 42.348527), zoom=8, maptype = "toner-lite", source = "stamen")

```

### Correct lat long

```{r eval=F}

BOS_basemap <- get_stamenmap(bbox = c(left = -73,
                                    bottom = 41,
                                    right = -69,
                                    top = 43.75), 
                             maptype = "toner-lite", source = "stamen")

#xlim = c(-73, -69), ylim = c(41, 43.75)

```

### Toner Basemap (correct)

```{r eval=F}

BOS_basemap_toner <- get_stamenmap(bbox = c(left = -73,
                                    bottom = 41,
                                    right = -69,
                                    top = 43.75), 
                             maptype = "toner", source = "stamen")

#xlim = c(-73, -69), ylim = c(41, 43.75)

```

### Google

```{r eval=F}

BOS_basemap <- get_map(location=c(lon = -71.110049, lat = 42.348527), zoom=8, maptype = "roadmap", source = "google")

```

### Didn't work

```{r eval=F}

BOS_basemap <- get_map(location=c(lon = -71.110049, lat = 42.348527), zoom=8, maptype = "roadmap")

```

## Filter sf

```{r}

BOS_data <- CSA_map %>% filter(QCEW_CSA == "CS148")

```


## ggplot

### Example Color Code:

scale_fill_distiller(palette = "YlGn")
scale_fill_viridis_c(alpha = .4)

From A Dictionary of Color Combinations:
https://sanzo-wada.dmbk.io/
scale_fill_gradient(low = "#b3d9a3", high = "#76844E")
scale_fill_gradient(low = "#e2ece7", high = "#9FC2B2")

### No Basemap

```{r}

BOS_map_EmpPct <- ggplot(BOS_data) +
  geom_sf(aes(fill = emp_CSAp))+
  scale_fill_gradient(low = "#d5dac9", high = "#76844E")

BOS_map_EmpPct

```

### Basemap

"#d5dac9"

```{r}

BOS_basemap_EmpPct <- ggmap(BOS_basemap) +
  geom_sf(data = BOS_data, aes(fill = emp_CSAp), inherit.aes = FALSE, alpha = .4) +
  scale_fill_gradient(low = "#d5dac9", high = "#76844E") +
  scale_alpha(.4) + 
  coord_sf(crs = st_crs(4326))

BOS_basemap_EmpPct

```

#### Export

```{r eval = F}

ggsave("BOS_emp_basemap.png", plot = BOS_basemap_EmpPct)

```


# osmdata Map

https://ggplot2tutor.com/tutorials/streetmaps
http://joshuamccrain.com/tutorials/maps/streets_tutorial.html

https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html#1_Introduction


## Import Features

WARNING: This is very slow, and creates very big files!

```{r eval=F}

library(osmdata)

streets <- opq(bbox = c(-72.251587,41.323201,-69.955444,43.068888)) %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary", 
                            "secondary", "tertiary")) %>%
  osmdata_sf()
small_streets <- opq(bbox = c(-72.251587,41.323201,-69.955444,43.068888)) %>%
  add_osm_feature(key = "highway", 
                  value = c("residential", "living_street",
                            "unclassified",
                            "service", "footway")) %>%
  osmdata_sf()

river <- opq(bbox = c(-72.251587,41.323201,-69.955444,43.068888)) %>%
  add_osm_feature(key = "waterway", value = "river") %>%
  osmdata_sf()

```

### Reset Overpass server

This one seems better: https://wiki.openstreetmap.org/wiki/Overpass_API#Public_Overpass_API_instances

```{r}

new_url <- "https://lz4.overpass-api.de/api/interpreter"
set_overpass_url(new_url) # reset the base url (not run)

```

### Much less!

xlim = c(-73, -69), ylim = c(41, 43.75)

getbb("Boston")%>%

opq(bbox = c(-73, 41, -69, 43.75))%>%

```{r}

streets <- getbb("Boston")%>%
  opq() %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway")) %>%
  osmdata_sf()

```


## Plot Map

```{r}

ggplot() +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .4,
          alpha = .8) +
  geom_sf(data = small_streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .4,
          alpha = .6) +
  geom_sf(data = river$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .2,
          alpha = .5) +
  coord_sf(xlim = c(7.77, 7.92), 
           ylim = c(47.94, 48.06),
           expand = FALSE) 

```

# Simple Geography

https://r-spatial.org/r/2018/10/25/ggplot2-sf-2.html 

```{r eval=FALSE}

library("rnaturalearth")
library("rnaturalearthdata")
library("maps")

world <- ne_countries(scale = "medium", returnclass = "sf")
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))

```

## NOTE

The "world" object doesn't match up with lines from "maps" package

```{r eval=FALSE}

BOS_simple_EmpPct <- ggplot(data = world) +
  geom_sf() +
  geom_sf(data = BOS_data, aes(fill = emp_CSAp), inherit.aes = FALSE, alpha = .4) +
  geom_sf(data = counties, fill = NA, color = gray(.5)) +
  geom_sf(data = states, fill = NA) + 
  scale_fill_gradient(low = "#d5dac9", high = "#76844E") +
  coord_sf(xlim = c(-73, -69), ylim = c(41, 43.75), expand = FALSE)

BOS_simple_EmpPct

```

# Using Census Shapefiles

```{r eval=F}

library(tidycensus)

counties_df <- get_decennial(geography = "county",
                       variables = "P001001",
                       geometry = T,
                       output = "wide")

```

```{r eval=FALSE}

BOS_census_EstPct <- ggplot(data = counties_df) +
  geom_sf(fill = "gray", color = gray(.5)) +
  geom_sf(data = BOS_data, aes(fill = est_CSAp), inherit.aes = FALSE, alpha = .4) +
  #geom_sf(data = counties, fill = NA, color = gray(.5)) +
  #geom_sf(data = states, fill = NA) + 
  scale_fill_gradient(low = "#e2ece7", high = "#9FC2B2") +
  coord_sf(xlim = c(-73, -69), ylim = c(41, 43.75), expand = FALSE)

BOS_census_EstPct

```

### Logged Employment

I logged employment to try and control for Middlesex County

NOTE: geom_sf(data = BOS_data, aes(fill = *log(lfSc_mp))*, inherit.aes = FALSE, alpha = .4)

```{r eval=FALSE}

BOS_census_emp <- ggplot(data = counties_df) +
  geom_sf(fill = "gray", color = gray(.5)) +
  geom_sf(data = BOS_data, aes(fill = log(lfSc_mp)), inherit.aes = FALSE, alpha = .4) +
  #geom_sf(data = counties, fill = NA, color = gray(.5)) +
  #geom_sf(data = states, fill = NA) + 
  scale_fill_gradient(low = "#e2ece7", high = "#4f6159") +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "white",
          size = .4,
          alpha = .4) +
  coord_sf(xlim = c(-73, -69), ylim = c(41, 43.75), expand = FALSE)

BOS_census_emp

```

#### Export

```{r eval = F}

ggsave("BOS_emp_roads.png", plot = BOS_census_emp)

```


# Geography Highlighted

## Filter sf

```{r}

BOS_UWS_data <- CSA_map %>% filter(are_fps == "25025" | are_fps == "25017")

BOS_suffolk_data <- CSA_map %>% filter(are_fps == "25025")

BOS_middlesex_data <- CSA_map %>% filter(are_fps == "25017")

```

## Map

```{r eval=FALSE}

BOS_UWS_counties <- ggplot(data = counties_df) +
  geom_sf(fill = "gray", color = gray(.5)) +
  geom_sf(data = BOS_suffolk_data, fill = "#9FC2B2", inherit.aes = FALSE, alpha = .8) +
  geom_sf(data = BOS_middlesex_data, fill = "#e2ece7", inherit.aes = FALSE, alpha = .8) +
  #geom_sf(data = counties, fill = NA, color = gray(.5)) +
  #geom_sf(data = states, fill = NA) + 
  #scale_fill_gradient(low = "#e2ece7", high = "#9FC2B2") +
  coord_sf(xlim = c(-73, -69), ylim = c(41, 43.75), expand = FALSE) +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "white",
          size = .4,
          alpha = .4) +
  coord_sf(xlim = c(-73, -69), ylim = c(41, 43.75), expand = FALSE)

BOS_UWS_counties

```

#### Export

```{r eval = F}

ggsave("BOS_UWS_roads.png", plot = BOS_UWS_counties)

```
# Zoomed In

## Roads Map

```{r eval=FALSE}

BOS_UWS_counties_zoom <- ggplot(data = counties_df) +
  geom_sf(fill = "gray", color = gray(.5)) +
  geom_sf(data = BOS_suffolk_data, fill = "#9FC2B2", inherit.aes = FALSE, alpha = .8) +
  geom_sf(data = BOS_middlesex_data, fill = "#e2ece7", inherit.aes = FALSE, alpha = .8) +
  #geom_sf(data = counties, fill = NA, color = gray(.5)) +
  #geom_sf(data = states, fill = NA) + 
  #scale_fill_gradient(low = "#e2ece7", high = "#9FC2B2") +
  coord_sf(xlim = c(-73, -69), ylim = c(41, 43.75), expand = FALSE) +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "white",
          size = .4,
          alpha = .4) +
  coord_sf(xlim = c(-72, -70), ylim = c(41.5, 43), expand = FALSE)

BOS_UWS_counties_zoom

```

#### Export

```{r eval = F}

ggsave("BOS_UWS_roads_zoom.png", plot = BOS_UWS_counties_zoom)

```

### Basemap

"#d5dac9"

```{r}

BOS_basemap_EmpPct_zoom <- ggmap(BOS_basemap) +
  geom_sf(data = BOS_suffolk_data, fill = "#9FC2B2", inherit.aes = FALSE, alpha = .4) +
  geom_sf(data = BOS_middlesex_data, fill = "#e2ece7", inherit.aes = FALSE, alpha = .4) +
  scale_fill_gradient(low = "#d5dac9", high = "#76844E") +
  scale_alpha(.4) + 
  coord_sf(xlim = c(-72, -70.5), ylim = c(42.1, 42.8), expand = FALSE)

BOS_basemap_EmpPct_zoom

```

#### Export

```{r eval = F}

ggsave("BOS_UWS_basemap_zoom.png", plot = BOS_basemap_EmpPct_zoom)

```
# osm second try

```{r eval = F}

library(osmdata)

streets <- opq(bbox = c(-72.251587,41.323201,-69.955444,43.068888)) %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway")) %>%
  osmdata_sf()

```

